\documentclass{amsart}
\usepackage{amssymb}
\renewcommand{\baselinestretch}{1.5}
\addtolength{\textwidth}{.2in}
\addtolength{\topmargin}{-.5in}
\addtolength{\textheight}{1in}

\usepackage{ifthen}

\input{my_latex_definitions.tex}

\addtolength{\abovedisplayskip}{0pt}
\addtolength{\belowdisplayskip}{24pt}
\addtolength{\abovedisplayshortskip}{0pt}
\addtolength{\belowdisplayshortskip}{44pt}


\begin{document}
\thispagestyle{empty}

\centerline{\Large Activity 16 -- Introduction to Proof}
\centerline{\large indirect arguments}

\bigskip
\Large


\begin{enumerate}

\item Prior to starting a proof by contradiction you must first determine the logical negation of the statement you are hoping to prove.  Find negations of the following.
\begin{enumerate}
\item \rule[-6pt]{0pt}{24pt} There are infinitely-many natural number $p$ such that $p$ is prime and $p+2$ is prime.
\item \rule[-6pt]{0pt}{24pt} $\displaystyle \forall n \in \Naturals, \; n^3 \; \mbox{is even} \; \implies \; n \; \mbox{is even} $
\item \rule[-6pt]{0pt}{24pt} $\displaystyle \forall n \in \Naturals, \; ( n \geq 2 ) \; \implies (\exists p,q \in \mbox{Primes}, \; p+q \, = \, 2n$
\end{enumerate}

\vfill

\item Proofs by contradiction and by contraposition are considered equivalent because, if you have a proof of one sort, you can recast it as the other sort.  If we are trying to prove a universal conditional sentence $\forall x, P(x) \implies Q(x)$,
a proof by contradiction consists in showing that $\exists x, (P(x) \land \lnot Q(x)) \implies c$, while a proof by contraposition consists of showing that  $\forall x, \lnot Q(x) \implies \lnot P(x)$.

Use a truth table to determine whether $(P \land \lnot Q) \implies c$ and $\lnot Q \implies \lnot P$ are equivalent.

\vfill

\vfill

\newpage

\item Devise a proof by contraposition and a proof by contradiction for item (b) in problem 1.

\vfill

\item Show, by contradiction, that there is no smallest positive real number.

\vfill

\newpage

\item Show that $\displaystyle \forall x, y \in \Reals, (x \in \Rationals \; \land \; xy \notin \Rationals) \; \implies \; y \notin \Rationals$.

\vfill

\item Prove that $\displaystyle \forall x, y \in \Integers, \; (x >1 ) \; \implies \; ( x \nmid y \; \lor \; x \nmid y+1 )$

\vfill

\newpage

\item Let's do one that's not an indirect argument.  (Just to mix things up a little.) The squares of natural numbers can only have certain values mod 4.  In particular, $\forall x \in \Naturals, \; x^2 \mod{4} = 0 \; \mbox{or} \; x^2 \mod{4} = 1$.  You can prove this by looking at two possibilities: either $x$ is even, or $x$ is odd.

\vfill

\item A \emph{Pythagorean triple} is a set of three
natural numbers, $a$, $b$ and $c$, that satisfy the Pythagorean theorem: $a^2 + b^2 = c^2$.   Prove that, in a
Pythagorean triple, at least one of $a$ and $b$ is even. (Use contradiction -- it will involve the statement we just proved in problem 7.

\vfill

\end{enumerate}

\end{document}
