\documentclass{amsart}
\usepackage{amssymb}
\renewcommand{\baselinestretch}{1.5}
\addtolength{\textwidth}{.2in}
\addtolength{\topmargin}{-.5in}
\addtolength{\textheight}{1in}

\usepackage{ifthen}

\input{my_latex_definitions.tex}

\addtolength{\abovedisplayskip}{0pt}
\addtolength{\belowdisplayskip}{24pt}
\addtolength{\abovedisplayshortskip}{0pt}
\addtolength{\belowdisplayshortskip}{44pt}


\begin{document}
\thispagestyle{empty}

\centerline{\Large Activity 19 -- Introduction to Proof}
\centerline{\large proofs and disproofs of existentials}

\bigskip
\Large


\begin{enumerate}

\item Is there a Fibonacci number that is a perfect cube?

\vfill

\item Partition numbers count the number of ways we can write a number $n$ as a sum (we don't count different orderings of the same sum).   We denote the partition numbers by $p(n)$.  By convention, $p(0)$ is defined to be 1 ($p(1)$ is also 1).  As a more interesting example, $p(4) = 5$ since there are 5 distinct ways to write 4 as a sum: $4$, $3+1$, $2+2$, $2+1+1$ and $1+1+1+1$.  The partition numbers are A000041 in the OEIS.

After the first two 1's the values of $p(n)$ are all prime (at first).  Prove that there is a non-prime partition number, $p(n)$ with $n \geq 2$.

\vfill

\item Show that there is a 4 digit vampire number.

\vfill

\newpage

\centerline{\bf Review problems for chapter 3.}

\item Use proof by exhaustion to show that there are no 2 digit vampire numbers.

\vfill

\item Prove
\[ \forall a, b, c \in \Integers, \; ( a \mid b \, \land \, b \mid c ) \implies a \mid c. \]

\vfill

\newpage

\item Prove 
\[ \forall a, b \in \Integers, \; ( a \mid b \, \land \, b \mid a ) \implies a=b. \]

\vfill 

\item Prove 
\[ \forall x \in \Reals, \; x^5 < x^4 \; \implies 8 > 5x+3. \]

\vfill

\newpage

\item Prove that for all integers $a, b,$ and $c$, if $a \mid b$ and $a \mid (b+c)$, then
$a|c$.

\vfill

\item Show that $\binom{n}{k} \cdot \binom{k}{r} \; = \; \binom{n}{r} \cdot \binom{n-r}{k-r}$ (for all integers $r$, $k$ and $n$ with $r \leq k \leq n$). 

\vfill

\newpage

\item Referring to the partition numbers defined in problem 2, we say a partition has ``odd parts'' if all the numbers in the sum are odd.  We say a partition has ``distinct parts'' if all the numbers in the sum are different from one another.  It is a general fact that the number of partitions of $n$ into odd parts is equal to the number of partitions of $n$ into distinct parts. 
(Frankly, that is really weird\textellipsis)

Show exhaustively that the number of partitions of 7 into odd parts is equal to the number of partitions of 7 into distinct parts.

\vfill

\end{enumerate}

\end{document}
