\documentclass{amsart}
\usepackage{amssymb}
\renewcommand{\baselinestretch}{1.5}
\addtolength{\textwidth}{.2in}
\addtolength{\topmargin}{-.5in}
\addtolength{\textheight}{1in}

\usepackage{ifthen}

\input{my_latex_definitions.tex}

\addtolength{\abovedisplayskip}{0pt}
\addtolength{\belowdisplayskip}{24pt}
\addtolength{\abovedisplayshortskip}{0pt}
\addtolength{\belowdisplayshortskip}{44pt}


\begin{document}
\thispagestyle{empty}

\centerline{\Large Activity 21 -- Introduction to Proof}
\centerline{\large containment}

\bigskip
\Large


\begin{enumerate}

\item Name three elements of $\Naturals$.  

\vfill

\item Name three subsets of $\Naturals$. 

\vfill

\item The empty set, $\emptyset$ is contained in every set $A$ (regardless of $A$'s contents).  Why?

\vfill

\item Let $A$ be the set $\{ \emptyset, 1, 2, \{3, 4, 5\} \}$.
Mark the following statements True or False.

\begin{tabular}{lcl}
\rule{0pt}{60pt}a)  $\emptyset \subseteq A$ & \rule{72pt}{0pt} & b)  $\emptyset \in A$ \\
\rule{0pt}{60pt}c)  $ \{1\} \subseteq A$ & & d) $ \{3, 4, 5\}  \subseteq A$ \\
\rule[-60pt]{0pt}{120pt}e)  $ \{ \emptyset \} \subseteq A$ & & f) $\{ 1, 2 \} \in A$\\
\end{tabular}

\newpage

\item Insert either $\in$ or $\subseteq$ in the blanks in the following 
sentences (in order to produce true sentences).

\vspace{.1in}

\begin{tabular}{lcl}
\rule{0pt}{16pt}i) $1$ \underline{\rule{36pt}{0pt}} $\{3, 2, 1, \{a, b\}\}$ & \rule{36pt}{0pt} & iii) $\{a, b\}$  \underline{\rule{36pt}{0pt}} $\{3, 2, 1, \{a, b\}\}$ \\
\rule{0pt}{16pt}ii) $\{a\}$ \underline{\rule{36pt}{0pt}} $\{a, \{a, b\}\}$ & &
iv) $\{\{a, b\}\}$  \underline{\rule{36pt}{0pt}} $\{a, \{a, b\}\}$ \\
\end{tabular}

\vspace{.2in}


\item A number is called {\em doubly even} if it is divisible by $4$.  Let $D$ denote the set of doubly-even natural numbers, and let $E$ denote the set of even natural numbers.  Which set is a subset of the other?

\vfill

\item The membership criteria for the sets $D$ and $E$ in the previous question are $M_D = 4 \divides x$ and $M_E = 2 \divides x$ respectively.  Which logical statement implies the other?

\vfill

\item The following incomplete diagram illustrates the correspondence between Set Theory (in the top row) and Logic (in the bottom row).  Fill in the missing symbols.

\begin{center}
\begin{tabular}{ccc}
\rule[-12pt]{0pt}{36pt} \rule{18pt}{0pt} $D$ \rule{18pt}{0pt} & \rule{18pt}{0pt} $\subseteq$ \rule{18pt}{0pt} & \rule{40pt}{0pt}  \\
\rule[-12pt]{0pt}{36pt}$\updownarrow$ & $\updownarrow$ & $\updownarrow$ \\
\rule[-12pt]{0pt}{36pt}  &    & $M_E(x)$ 
\end{tabular}
\end{center}

\vspace{.5in}

\newpage

\item Name a superset of $\Naturals$.

\vfill

\item Name a superset of $\Reals$.

\vfill

\item In the next lecture we'll talk about set equality.  How would you describe a reasonable way to determine if two sets are ``the same''?

\vfill

\end{enumerate}

\end{document}
