\documentclass{amsart}
\pdfminorversion=7 
\usepackage{amssymb}
\renewcommand{\baselinestretch}{1.5}
\addtolength{\textwidth}{.2in}
\addtolength{\topmargin}{-.5in}
\addtolength{\textheight}{1in}

\usepackage{ifthen}
\usepackage{graphicx}
\usepackage{color}

\input{my_latex_definitions.tex}

\addtolength{\abovedisplayskip}{0pt}
\addtolength{\belowdisplayskip}{24pt}
\addtolength{\abovedisplayshortskip}{0pt}
\addtolength{\belowdisplayshortskip}{48pt}
\newcommand{\vsp}{\rule[-12pt]{0pt}{48pt}}

\begin{document}
\thispagestyle{empty}

\centerline{\Large Activity 31 -- Introduction to Proof}
\centerline{\large equivalence relations}

\bigskip
\Large


\begin{enumerate}
	
\item (Exercise 1 in GIAM section 6.3) Consider the relation $\relA$ defined by 
\[ \relA = \{ (x,y) \suchthat \; x \, \mbox{has the same astrological sign as} \, y \}. \]


Verify that $\relA$ is reflexive, symmetric, and transitive.

\vfill

\vfill

\newpage

\item Consider the generic ``equivalence mod $m$'' relation:

\[ x \relR y \quad \iff \quad x \mod{m} \; = \; y \mod{m} . \]

Show that $\relR$ is reflexive, symmetric and transitive.  

\vfill

\newpage


\item In the book we defined the function $sf(n)$ which returns $n$ divided by the largest perfect square that divides $n$.  For example $sf(20) = 5$ since $4$ is the largest perfect square that divides $20$, and $20/4 = 5$ 

Use the $sf$ function to define a relation:

\[ x \relS y \quad \iff \quad sf(x) = sf(y) \]

Show that $\relS$ is an equivalence relation on $\Naturals$.

\vfill 

\newpage

\item Continuing with the relation $\relS$ defined in problem 4, notice that for any number $x$ that actually is a perfect square, $sf(x) = 1$.  Given this, what is $\overline{1}$?

\vfill


\item (Still working with the relation from problem 4.)  Many different natural numbers can be used as a ``label'' for the equivalence classes under $\relS$.  For instance $\overline{2}$, $\overline{8}$ and $\overline{18}$ all denote the same equivalence class -- let's follow the convention of always using the smallest such label.  Characterize these reduced representatives for the equivalence classes.

\vfill

\newpage

\item Define a relation $\relQ$ on the set of all finite sets by

\[ A \relQ B \quad \iff \quad  |A| = |B| \]

What are the equivalence classes under $\relQ$?

\vfill

\item (Exercise 3 in GIAM section 6.3.)

Define a relation $\relA$ on the set of all words by

\[ w_1 \relA w_2 \quad \iff \quad w_1 \mbox{ is an anagram of } w_2. \]

\noindent Show that $\relA$ is an equivalence relation.  (Words are anagrams
if the letters of one can be re-arranged to form the other.  For example, `ART' and `RAT' are anagrams.)

\vfill

\newpage

\item (Exercise 4 in GIAM section 6.3.)

The two diagrams below both show a famous graph known as the 
\index{Petersen graph}Petersen graph.  The picture on the 
left is the usual representation which emphasizes its five-fold symmetry.  The picture on the right
highlights the fact that the Petersen graph also has a three-fold symmetry.  Label the right-hand diagram
using the same letters (A through J) in order to show that these two representations are truly isomorphic.

\vspace{.2in}

\rule{0pt}{0pt} \hspace{-.75in} \input{petersen_iso.tex}

\vspace{.2in}

\vfill

\end{enumerate}


\end{document}
