\documentclass{amsart}
\usepackage{amssymb}
\renewcommand{\baselinestretch}{1.5}
\addtolength{\textwidth}{.2in}
\addtolength{\topmargin}{-.5in}
\addtolength{\textheight}{1in}

\usepackage{ifthen}
\usepackage{graphicx}

\input{my_latex_definitions.tex}

\addtolength{\abovedisplayskip}{0pt}
\addtolength{\belowdisplayskip}{24pt}
\addtolength{\abovedisplayshortskip}{0pt}
\addtolength{\belowdisplayshortskip}{48pt}

\newcommand{\vsp}{\rule[-12pt]{0pt}{48pt}}

\begin{document}
\thispagestyle{empty}

\centerline{\Large Activity 30 -- Introduction to Proof}
\centerline{\large properties of relations}

\bigskip
\Large


\begin{enumerate}
	
\item Name a relation that is irreflexive, anti-symmetric, and transitive.

\vfill

\item (Exercise \# 4 in Section 6.2 of GIAM) For each of the five properties, name a relation that has it and a relation that doesn’t.

\vfill

\item Consider the relation $\relR$ defined by 

\[ x \relR y \quad \iff \quad x \pmod{3} \; = \; y \pmod{3} . \]

Show that $\relR$ is reflexive.

\vfill

\newpage

\item Show that $\relR$ from the previous problem is symmetric.

\vfill

\item To show that $\relR$ (defined in problem 2) is transitive we need to assume that there are 3 arbitrary integers, $a$, $b$, and $c$ such that $a \relR b$ and $b \relR c$.  From that setup we must deduce that $a \relR c$.

Use the definition of $\relR$ to write down precisely what needs to be shown to prove that this relation is transitive.

\vfill 

\newpage

\item Critique the following proof that $\relR$ is transitive.

\begin{proof} Suppose that $a$, $b$ and $c$ are integers such that 
 $a \pmod{3} \; = \; b \pmod{3}$ and $ b \pmod{3} \; = \; c \pmod{3}$.  It follows that $ a \pmod{3} \; = \; c \pmod{3}$ because of ``reasons.''
\end{proof}

\vfill


\item In the last worksheet we defined a relation on $\Reals$ by $x \relR y \; \iff x^2 - y^2 \; = \; 0$.  Verify that this relation is reflexive and symmetric by looking at its graph.  Prove that it is also transitive. 

\vfill

\newpage

\item Define a relation $L$ on the set of all words in the English language via

\[ xLy \quad \iff \quad x \; \mbox{comes before} \; y \; \mbox{in the dictionary}. \]

Which of the 5 properties does $L$ have?

\vfill

\item Let $S$ be the power set of $\{1, 2, 3 \}$ and consider the $\subseteq$ relation on $S$.  Informally verify that this is an ordering relation.  (Which three properties must we check?)


\vfill

\end{enumerate}


\end{document}
