\ifx\pdfoutput\undefined % We're running regular LaTeX (not pdftex)
\documentclass[dvips,12pt,twoside]{book}
\RequirePackage[plainpages=false]{hyperref}
\usepackage{color}
\typeout{Use 'fig4latex -l' then 'make' to format graphics for regular latex}
\else % We're running PdfTeX
\documentclass[pdftex,12pt,twoside]{book}
%\usepackage{thumbpdf}
\RequirePackage[pdftex]{hyperref}
\hypersetup{colorlinks=true, %
  plainpages=false, %
  pdfpagelabels, %
  %hyperindex=true, %
  %backref=true, %
  %bookmarks=true, %
  pdftitle={A Gentle Introduction to the Art of Mathematics, Version 3.1}, %
  pdfauthor={Joseph E. Fields}, %
  pdfsubject={Foundations of Mathematics}, %
  %pdfpagelayout=SinglePage, %
  %pdfpagetransition=Dissolve, %
  pdfstartview=FitH, %
  %pdfstartpage=1, %
  %pdffitwindow=true
}
\pdfcompresslevel=9
\typeout{Use 'fig4latex -p' then 'make' to format graphics for pdflatex.}

\fi

\usepackage[english]{babel} %language selection
\selectlanguage{english}

\usepackage{makeidx}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{amssymb}

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{color}

\usepackage{ifthen}

%These booleans allow us to generate 3 different books from the same LaTeX source

\newboolean{InTextBook}
\setboolean{InTextBook}{true}
\newboolean{InWorkBook}
\setboolean{InWorkBook}{false}
\newboolean{InHints}
\setboolean{InHints}{false}

%When this boolean is true (beginning in Section 5.1) we will use the convention
% that $0 \in \Naturals$.  If it is false we will continue to count $1$ as the smallest
%natural number (thus making Giuseppe Peano spin in his grave...)
 
\newboolean{ZeroInNaturals}

%This boolean is used to distinguish the version where we use $\sim$ rather than $\lnot$

\newboolean{LNotIsSim}

%The values of the last two booleans are set in ``switches.tex''

\input{switches}

\let\savedlnot\lnot
\ifthenelse{\boolean{LNotIsSim}}{\renewcommand{\lnot}{\sim} }{}

%This command puts different amounts of space depending on whether we are
% in the text, the workbook or the hints & solutions manual. 
\newcommand{\twsvspace}[3]{%
 \ifthenelse{\boolean{InTextBook} }{\vspace{#1}}{%
  \ifthenelse{\boolean{InWorkBook} }{\vspace{#2}}{%
   \ifthenelse{\boolean{InHints} }{\vspace{#3}}{} %
   }%
  }%
 }

\newcommand{\wbvfill}{\ifthenelse{\boolean{InWorkBook}}{\vfill}{}}
\newcommand{\textbookpagebreak}{\ifthenelse{\boolean{InTextBook}}{\newpage}{}}
\newcommand{\workbookpagebreak}{\ifthenelse{\boolean{InWorkBook}}{\newpage}{}}
\newcommand{\hintspagebreak}{\ifthenelse{\boolean{InHints}}{\newpage}{}}

\newcommand{\hint}[1]{\ifthenelse{\boolean{InHints}}{{ \color[rgb]{0,0,1} #1} }{}}

%All my customized LaTeX commands are in this file:

\input{my_latex_definitions.tex}

%Things that effect the size and placement of text on the page

\renewcommand{\baselinestretch}{1.3}
\renewcommand{\arraystretch}{.77}

\addtolength{\textheight}{.25in}
\addtolength{\oddsidemargin}{.75in}
\addtolength{\evensidemargin}{-.75in}

\makeindex

\begin{document}

\frontmatter

\title{A Gentle Introduction to the Art of Mathematics\\ {\small Version 3.1 \ifthenelse{\boolean{LNotIsSim}}{S}{}\ifthenelse{\boolean{ZeroInNaturals}}{}{N} }}
\author{Joseph Fields}
\date{Southern Connecticut State University}

\maketitle

\clearpage

\rule{0pt}{0pt}

\vfill

\begin{quote}
    Copyright \copyright{}  2013  Joseph E. Fields.
    Permission is granted to copy, distribute and/or modify this document
    under the terms of the GNU Free Documentation License, Version 1.3
    or any later version published by the Free Software Foundation;
    with no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts.
    A copy of the license is included in the section entitled ``GNU
    Free Documentation License''.
\end{quote}

\vfill

\begin{quote}
The latest version of this book is available (without charge) in portable document format at \newline
\rule{0pt}{0pt} \hspace{1in} \verb+http://www.southernct.edu/~fields/+
\end{quote}

\vfill

\clearpage

\rule{0pt}{0pt}

\vfill

\begin{quote}
{\Large \bf Acknowledgments} 

   This is version 3.1 of {\em A Gentle Introduction to the Art of Mathematics}.
Earlier versions were used and classroom tested by several colleagues: 
Robert Vaden-Goad, John Kavanagh, Ross Gingrich, Aaron Clark.  I thank you all.  
A particular debt of gratitude is owed to Len Brin whose keen eyes caught 
a number of errors and inconsistencies, and who contributed many new 
exercises.  Thanks, Len. 
  
\end{quote}

\vfill


\begin{quote}
{\Large \bf A word about options} 

There are currently 4 optional versions of this text.  The different options are distinguished with
flags following the version number.  Thus you may see a plain ``version  3.1'' (no flags) 
or any of the following: ``version 3.1S'', ``version 3.1N'',  or ``version 3.1SN''. 
The S flag is for those who prefer the symbol $\sim$ to represent 
logical negation.  (The default is to
use the symbol $\savedlnot$ for logical negation.)  The N flag distinguishes a version in which
the convention that $1$ is the smallest natural number is followed throughout.  The absence
of the N flag indicates the original version of the text in which the naturals do not contain $0$
in chapters 1 through 4 and (after the Peano axioms are introduced) that convention is changed
to the more modern rule. 

\end{quote}


\vfill
\clearpage

\tableofcontents

\listoffigures

\listoftables

\input{pref1}
\input{pref2}

\mainmatter

\include{intro/intro}
\include{logic/logic}
\include{proof1/proof1}
\include{sets/sets}
\include{proof2/proof2}
\include{rel/rel}
\include{proof3/proof3}
\include{card/card}
\include{proof4/proof4}

\bibliographystyle{plain}
\bibliography{main}
\addcontentsline{toc}{chapter}{References}
\phantomsection
\markboth{REFERENCES}{REFERENCES}%

%\addcontentsline{toc}{chapter}{GNU Free Documentation License}
%\phantomsection 
%\markboth{GFDL}{GFDL}%

%\include{fdl-1.3}

\addcontentsline{toc}{chapter}{Index} 
\phantomsection
\markboth{INDEX}{INDEX}
\printindex

\end{document}

