\begin{enumerate}

\item The transitive property of equality says that if $a=b$ and $b=c$
then $a=c$.  Does the implication arrow satisfy a transitive property?
If so, state it.

\item Complete truth tables for the compound sentences $A \implies B$ and
  ${\lnot}A \lor B$.

\item Complete a truth table for the compound sentence $A \implies (B \implies C)$ and for the sentence $(A \implies B) \implies C$.  What can you conclude
about conditionals and the associative property?

\item Determine a sentence using the {\em and} connector ($\land$) that
gives the negation of $A \implies B$.

\item Rewrite the sentence ``Fix the toilet or I won't pay the rent!'' as
a conditional.

\item Why is it that the sentence ``If pigs can fly, I am the king
of Mesopotamia.'' true?

\item Express the statement $A \implies B$ using the Peirce arrow and/or the
Scheffer stroke. (See Exercise~\ref{ex:nand_nor} in the previous section.)

\item Find the contrapositives of the following sentences.
  \begin{enumerate}
  \item If you can't do the time, don't do the crime.
  \item If you do well in school, you'll get a good job.
  \item If you wish others to treat you in a certain way, you must 
    treat others in that fashion.
  \item If it's raining, there must be clouds.
  \item If $a_n \leq b_n$, for all $n$ and $\sum_{n=0}^\infty b_n$ is a 
convergent series, then $\sum_{n=0}^\infty a_n$ is a convergent series.
  \end{enumerate}

\item What are the converse and inverse of ``If you watch my back, I'll 
watch your back.''?

\item The integral test in Calculus is used to determine whether an
infinite series converges or diverges:   Suppose that $f(x)$ is a positive,
decreasing, 
real-valued function with $\lim_{x \longrightarrow \infty} f(x) = 0$, if
the improper integral
$\int_0^\infty f(x)$ has a finite value, then the infinite series 
$\sum_{n=1}^\infty f(n)$ converges.

The integral test should be envisioned by letting the series correspond
to a right-hand Riemann sum for the integral, since the function is decreasing,
a right-hand Riemann sum is an underestimate for the value of the integral,
thus

\[ \sum_{n=1}^\infty f(n) < \int_0^\infty f(x). \]

Discuss the meanings of and (where possible) provide justifications for
the inverse, converse and contrapositive of the conditional statement 
in the integral test.

\item On the Island of Knights and Knaves (see page~\pageref{IKK}) you encounter two individuals named Locke and Demosthenes.  

Locke says, ``Demosthenes is a knave.'' \newline
Demosthenes says ``Locke and I are knights.''

Who is a knight and who a knave?

\end{enumerate}
