\begin{enumerate}
\item Write an inductive proof of the formula for the sum 
of the first $n$ cubes.

\item Find a formula for the sum of the first $n$ fourth powers.

\item The sum of the first $n$ natural numbers is sometimes called
the $n$-th triangular number \index{triangular numbers}$T_n$.  Triangular numbers are so-named
because one can represent them with triangular shaped arrangements 
of dots. 

\begin{center} \input{figures/triangular_numbers.tex} \end{center}

The first several triangular numbers are 1, 3, 6, 10, 15, et cetera.

Determine a formula for the sum of the first $n$ triangular numbers $\displaystyle \left( \sum_{i=1}^n T_i \right)$ and prove it using PMI.

\item Consider the alternating sum of squares:
\begin{gather*}
1 \\
1 - 4 = -3 \\
1 - 4 + 9 = 6 \\
1 - 4 + 9 - 16 = -10 \\
\mbox{et cetera}
\end{gather*}
Guess a general formula for $\sum_{i=1}^n (-1)^{i-1} i^2$, and prove it using PMI.

\item Prove the following formula for a product.

\[ \prod_{i=2}^n \left(1 - \frac{1}{i}\right) =  \frac{1}{n} \]

\item Prove $\displaystyle \sum_{j=0}^{n}(4j+1) \; = \; 2n^{2}+3n+1$ for all
integers $n \geq 0$.

\item Prove $\displaystyle \sum_{i=1}^{n}\frac{1}{(2i-1)(2i+1)} \; = \; \frac{n}{2n+1}$ for all natural numbers $n$.

\item The \index{Fibonacci numbers} \emph{Fibonacci numbers} are a sequence of integers defined by
the rule that a number in the sequence is the sum of the two that 
precede it.

\[ F_{n+2} = F_n + F_{n+1}  \]

\noindent The first two Fibonacci numbers (actually the zeroth and the first) 
are both 1.  

\noindent Thus, the first several Fibonacci numbers are

\[ F_0 = 1, F_1=1, F_2=2, F_3=3, F_4=5, F_5=8, F_6=13, F_7=21, \; \mbox{et cetera} \]

Use mathematical induction to prove the following formula involving
Fibonacci numbers.

\[ \sum_{i=0}^n (F_i)^2 \, = \, F_n \cdot F_{n+1} \]

\end{enumerate}

%% Emacs customization
%% 
%% Local Variables: ***
%% TeX-master: "GIAM-hw.tex" ***
%% comment-column:0 ***
%% comment-start: "%% "  ***
%% comment-end:"***" ***
%% End: ***

