
\begin{enumerate}

\item For each of the following functions, give its domain, range and a possible codomain.
  \begin{enumerate}
  \item $f(x) = \sin{(x)}$
  \item $g(x) = e^x$
  \item $h(x) = x^2$
  \item $m(x) = \frac{x^2+1}{x^2-1}$
  \item $n(x) = \lfloor x \rfloor$
  \item $p(x) = \langle \cos{(x)}, \sin{(x)} \rangle $
  \end{enumerate}

\item Find a bijection from the set of odd squares, $\{1, 9, 25, 49, \ldots\}$,
to the non-negative integers, $\Znoneg = \{0,1,2,3, \ldots\}$.
Prove that the function you just determined is both injective and surjective.
Find the inverse function of the bijection above.

\item The natural logarithm function $\ln (x)$ is defined by a definite
integral with the variable $x$ in the upper limit.

\[ \ln (x) = \int_{t=1}^{x} \frac{1}{t} \, \mbox{d}t. \]

From this definition we can deduce that $\ln (x)$ is strictly increasing on its
entire domain, $(0, \infty)$.  Why is this true?

We can use the above definition with $x=2$ to find the value of 
$\ln (2) \approx .693$.  We will also take as given the following 
rule (which is valid for all logarithmic functions).

\[ \ln(a^b) = b \ln(a) \]

Use the above information to show that there is neither an upper bound 
nor a lower bound for the values of the natural logarithm.  These facts
together with the information that $\ln$ is strictly increasing show that
$\Rng{\ln} = \Reals$.
 
\item Georg Cantor developed a systematic way of listing the rational numbers.
By ``listing'' a set one is actually developing a bijection from $\Naturals$ to
that set.  The method known as ``Cantor's Snake'' creates a bijection from
the naturals to the non-negative rationals.  
First we create an infinite table whose rows
are indexed by positive integers and whose columns are indexed by non-negative
integers -- the entries in this table are rational numbers of the form
``column index'' / ``row index.''  We then follow a snake-like path that
zig-zags across this table -- whenever we encounter a rational number that 
we haven't seen before (in lower terms) we write it down.  This is indicated 
in the diagram below by circling the entries.   

Effectively this gives us a function $f$ which produces the rational number 
that would be found in a given position in this list.  For example 
$f(1) = 0/1, f(2) = 1/1$ and $f(5) = 1/3$.  

What is $f(26)$?  What is $f(30)$?  What is $f^{-1}(3/4)$? What is $f^{-1}(6/7)$?

\begin{center}
\input{figures/Cantor_snake.tex}
\end{center}
   
\end{enumerate}


%% Emacs customization
%% 
%% Local Variables: ***
%% TeX-master: "GIAM-hw.tex" ***
%% comment-column:0 ***
%% comment-start: "%% "  ***
%% comment-end:"***" ***
%% End: ***
